\section{Doyen Team Contract}

\begin{itemize}
    \item We will communicate primarily through Discord when not in person.
    \item We will meet on Mondays at 7:30pm ET, and otherwise as arranged as needed.
    \item Time zones:
    \begin{itemize}
        \item Nicholas - PST
        \item Edwin - EST
        \item Patrick - EST
        \item Jeremiah - PST
        \item Muhammad - EST
    \end{itemize}
    \item Before the team meeting, team members should:
    \begin{itemize}
        \item Contribute to the shared agenda
        \item Have questions, comments, or ideas related to the agenda
    \end{itemize}
    \item We will structure our work as follows:
    \begin{itemize}
        \item We will work asynchronously via GitHub and its projects/issues.
        \item We will break the work into week-long sprints with a customer demo each week.
        \item We will have a daily stand-up in Discord text, which answers the following questions:
        \begin{itemize}
            \item What did you do yesterday?
            \item What will you do today?
            \item Are there any impediments in your way?
        \end{itemize}
    \end{itemize}
    \item We will submit assignments as follows:
    \begin{itemize}
        \item The team works independently for individual submissions.
        \item The team will agree upon a due date for collaborative submission for each member to complete their portions before submission.
    \end{itemize}
    \item Surprises will be handled as follows:
    \begin{itemize}
        \item Daily stand-ups via Discord will inform the team as early as possible.
        \item The team will try not to build silos. We can try to overlap skills within the project’s subsystems.
    \end{itemize}
    \item Turn-taking will be managed as follows:
    \begin{itemize}
        \item We’ll make space at the end (and often the beginning) of each to give each team member space to speak and share their thoughts.
    \end{itemize}
    \item Conflict will be handled as follows:
    \begin{itemize}
        \item If it is a disagreement on our method, we will use a consensus process to try to resolve the disagreement.
        \item Speak with each other.
        \item Speak openly about struggles.
        \item Speak with TF as a neutral third party for a final fallback.
        \item Don’t take other people’s work without consent (within reason – someone could be very, very far behind).
    \end{itemize}
    \item Other considerations:
    \begin{itemize}
        \item Brainstorming / Design Critiques:
        \begin{itemize}
            \item Use building language and constructive criticism.
            \item Compliment sandwiches when needed.
        \end{itemize}
    \end{itemize}
\end{itemize}
