\section{Team Strategy and Dynamics}

We organized to tackle the project by discussing team members’ backgrounds and surveying them to gauge the skill set of each team member. The plan is to get contributions from every team member and keep the balance between learning and utilizing the existing skill set. This balance will be maintained through feedback and discussion.

We will be largely following an agile strategy for organizing our work and will make decisions by a consensus process if disagreements come up. We have set up a discord server to discuss project-related issues and get instant feedback from the team members. We have internal meetings twice a week and customer meetings once per week. We will meet each week, on Monday evenings and:
\begin{enumerate}
    \item Sprint Review
    \item Retrospective
    \item Review the Backlog
    \item Plan the next Sprint
\end{enumerate}
We will also have a standing meeting with the customer once weekly on Fridays and can meet more or less as needed. In summary:
\begin{itemize}
    \item Sprints - Weekly
    \item Roles - Flat structure, with specialization into aspects of the project as needed.
    \item Ceremonies - Daily standup (on Discord)
    \item Backlog refinement - Part of our weekly meetings
    \item Sprint Planning - Part of our weekly meetings
    \item Daily Scrum - Daily standup (on Discord)
    \item Sprint Review - Part of our weekly meetings
    \item Retrospectives - Part of our weekly meetings
\end{itemize}
